### 2018-07-30 paper/bibliography.tex
Visualisation working

John E. Hopcroft, Rajeev Motwani, Jeffrey D. Ullman, 'Introduction to Automata
Theory, Languages, and Computation', 3rd Edition, Prentice Hall, 2006

Hiroshi Umeo, Kinuo Nishide, Keisuke Kubo, 'A Simple Optimum-Time FSSP Algorithm
for Multi-Dimensional Cellular Automata', journal ref: EPTCS 90, 2012, pp.
151-165 (arXiv:1208.2761)

J. Gruska, S. La Torre, M. Napoli, M. Parente, 'Various Solutions to the Firing
Squad Synchronization Problems', 2005 (arXiv:cs/0511044)

Danny Dolev, Ezra N. Hoch, Yoram Moses, 'An Optimal Self-Stabilizing Firing
Squad', 2009 (arXiv:0908.2295)

Nazim Fatès, Irène Marcovici, Siamak Taati, 'Two-dimensional traffic rules and
the density classification problem', 2016 (arXiv:1604.04402)

Henryk Fuks, 'Solution of the Density Classification Problem with Two Cellular
Automata Rules', journal ref: Phys. Rev. E 55, 2081R (1997)
(arXiv:comp-gas/9703001)