### 2018-08-19 paper/paper.tex
More elementary CA. Updated paper

\documentclass[11pt,a4paper]{article}

\bibliographystyle{ieeetr}

\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{url}
\usepackage{diagbox}

\title{Cellular Automata and Computational Universality}

\begin{document}

\maketitle

\begin{center}
    Thomas Archbold \\
    University of Warwick \\
    \texttt{T.Archbold@warwick.ac.uk}
\end{center}

\begin{abstract}
    Cellular automata are discrete models with the ability to not only give rise
    to beautiful, intricate patterns, but also to be used as powerful tools of
    computation, with applications in cryptography, error-correction coding, and
    simulation of computer processors, to name a few. They also raise profound
    questions about the nature of our reality, asking whether our universe could
    be one such automaton. This paper provides a discussion into these automata,
    exploring the various power and limitations of a number of specific rule
    sets, covering John Conway's well-known ``Game of Life'' to the more obscure
    ``Langton's ant'' and ``Wireworld''. In particular, it explores the notion
    of computational universality, or Turing completeness, an automaton's
    ability to simulate any conceivable computation, and considers their
    potential in the context of solving two specific problems, the Firing Squad
    Synchronisation Problem and the Majority Problem. Existing solutions to
    these problems are explored, and their existing avenues for optimisation are
    discussed. In order to fully appreciate the complex structures that can
    arise from such simple beginnings, this project also presents software to
    visualise and probe further into the nature of the automata highlighted.
\end{abstract}

\section{Introduction}
    A cellular automaton is a discrete model of computation which consists of a
    finite collection of ``cells'', each in one of a finite number of states.
    The state of each of these cells may evolve over the progression of time in
    discrete steps, and may do so according to a deterministic set of rules
    based on the states of neighbouring cells. Their inherently discrete nature
    allows for strong analogies to be made with digital computers, and gifts
    them the ability to simulate digital processes and the potential to solve
    problems in this area.

    Consider the cellular automaton defined by the simple rule:
    \begin{equation}
        \label{eq:rule90}
        a_{i}^{t+1} = a_{i-1}^{t} + a_{i+1}^{t} \mod{2}
    \end{equation}

    For any automaton, in each time step the rule is applied to all cells in the
    automaton instantaneously and simultaneously. In this case, our set of
    states is ${0,1}$, and for each cell we look at the cells immediately
    preceding and succeeding it: if exactly one of them is in state 1, then this
    cell will be in state 1 in the next time step. Otherwise it will be in state
    0.

    \begin{figure}[h]
        \begin{center}
            \includegraphics[width=4in]{rule90.png}
            \caption{Pattern generated from applying Eq. 1 over 129 generations}
            \label{fig:rule90}
        \end{center}
    \end{figure}

    \subsection{Self-similarity}
    Figure \ref{fig:rule90} shows a visualisation of Eq. \ref{eq:rule90}. A cell is
    filled white if it has the value 1, and black otherwise. Note that each new
    ``line'' of the visualisation represents the automaton in the next
    successive time-step from the one before; this automaton is one-dimensional,
    and so we may represent it in its entirety as a single row of cells. This
    pattern exhibits two important characteristics, the first of which is
    ``self-similarity'', meaning that portions of the pattern generated are
    indistinguishable from the whole when magnified. As Wolfram states, in this
    way ``the pattern is therefore invariant under rescaling... and may be
    characterised by a fractal dimension'' \cite{WolframFractal}. One such
    definition of this is the Hausdorff-Besicovitch dimension \cite{Hausdorff}.

    As an example, suppose we have a cube with some mass, and we wish to find
    out how the mass scales when we try to make the same cube out of smaller
    copies of the original. Intuitively, one way to do this require eight
    smaller cubes, each of whose side length is half that of the original cube.
    So with the scaling factor of $\frac{1}{2}$, the mass of each smaller cube
    is $\frac{1}{8} = (\frac{1}{2})^3$.
    So we arrive at the equation $N = S^D$, where $N$ is the number of smaller
    copies that can be stuck together to make the original, $S$ is the scaling
    factor of these smaller copies, and $D$ is the dimension. In this example, we
    compute that the dimension of a cube is 3, which is obviously correct.  Back
    to the dimension of the Sierpinski Triangle, we can see that each larger
    triangle is made up of three smaller triangles, each of whose side lengths
    are half that of the larger triangle. So we calculate the dimension as:

    \begin{equation}
        \label{sierpinskiDim}
        \begin{split}
            \tfrac{1}{3} &= (\tfrac{1}{2})^D \\
            \log{\tfrac{1}{3}} &= D \log{\tfrac{1}{2}} \\
            D &= \tfrac{\log{3}}{\log{2}} \approx 1.585
        \end{split}
    \end{equation}

    Many naturally-occurring systems exhibit fractal structures, from snowflakes
    to pine cones to Romanesco broccoli, and raises the possibility that these
    are generated through the evolution of some natural cellular automata or
    similar processes.

    \begin{figure}[h]%
        \centering
        \subfloat[Snowflake]{{\includegraphics[width=5cm]{snowflake.jpg}}}%
        \qquad
        \subfloat[Romanesco broccoli]{{\includegraphics[width=5cm]{romanesco_broccoli.jpg}}}%
        \caption{Snowflakes and Romanesco broccoli exhibit fractal structures}%
        \label{fig:nature_fractals}%
    \end{figure}

    \subsection{Self-organisation}
    Second of the characteristics exhibited by this cellular automaton is
    ``self-organisation'', whereby from an initial disorderly or chaotic state
    there appear ordered structures seemingly spontaneously. This is illustrated
    in Fig. \ref{fig:self-organisation} below: one can clearly see the emergence
    of similar triangular structures in the cone snail's shell, appearing from a
    chaotic background. This starting chaos is modelled in the cellular
    automaton by randomly assigning each cell a value of 0 or 1.

    \begin{figure}[h]%
        \centering
        \subfloat[Shell of a cone
        snail]{{\includegraphics[width=2in]{cone_shell.jpg}}}%
        \qquad
        \subfloat[Rule 90 with random starting
        configuration]{{\includegraphics[width=2in]{rule90_rand.png}}}%
        \caption{Natural processes can exhibit structures found in automata}
        \label{fig:self-organisation}
    \end{figure}

\section{Classifying automata}
    In order to classify automata, we can take into account both the size of the
    neighbourhood each cell considers going into the next generation, as well as
    the number of possible states in which a cell may be. Call these $N$ and $Q$
    respectively; in the case of Eq. \ref{eq:rule90}, $N=3$ and $Q=2$, and so the
    total number of different ``transitions'' from one state to the next is $N^Q
    = 8$. We may write out all of these transitions as follows:

    \begin{table}[h]
        \centering
        \begin{tabular}{c|c|c|c|c|c|c|c}
            111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\
            \hline
            0   & 1   & 0   & 1   & 1   & 0   & 1   & 0
        \end{tabular}
    \end{table}

    Reading the bottom row of this transition table, we get the binary number
    $01011010_{\text{bin}} = 90_{\text{dec}}$, allowing us to name this rule set
    Rule 90. Using this convention, we may now construct our own automata based
    solely on the knowledge of its name (at least for elementary cellular
    automata). For example, take Rule 110, introduced by Wolfram 1983
    \cite{WolframRule110}. We first convert $110_{10}$ to binary, and then use
    this to fill in the table
    as above:

    \begin{table}[h]
        \begin{minipage}{.5\linewidth}
            \centering
            \begin{tabular}{c|c|c|c|c|c|c|c}
                111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\
                \hline
                ? & ? & ? & ? & ? & ? & ? & ?
            \end{tabular}
        \end{minipage}%
        $ \Rightarrow $
        \begin{minipage}{.6\linewidth}
            \begin{tabular}{c|c|c|c|c|c|c|c}
                111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\
                \hline
                0 & 1 & 1 & 0 & 1 & 1 & 1 & 0
            \end{tabular}
        \end{minipage}
    \end{table}

    To convert this to a concrete set of rules to simulate in a computer
    programs (in a more elegant way than a collection of eight \texttt{if}
    statements), we can put the information we have gathered so far in a
    Karnaugh map to get a Boolean algebraic expression of the ruleset, as in the
    table below. We use $A$ , $B$, and $C$ to denote $a_{i-1}$, $a_i$, and
    $a_{i+1}$, at time $t$, respectively. From this we can reduce the ruleset to
    the Boolean expression in Eq. \ref{eq:rule110} by observation.

    \begin{table}[h]
        %\caption{The Karnaugh map resulting from decoding Rule 110}
        %\label{kmap}
        \centering
        \begin{tabular}{|c||c|c|c|c|}\hline
            \diagbox{C}{AB} & 00 & 01 & 11 & 10 \\
            \hline
            0 & 0  & 1  & 1  & 0    \\
            \hline
            1 & 1  & 1  & 0  & 1    \\
            \hline
        \end{tabular}
    \end{table}

    \begin{equation}
        \label{eq:rule110}
        \begin{split}
            B^{t+1} &= \overline{A} \cdot C + B \cdot \overline{C} + \overline{B} \cdot C \\
            &= \overline{A} \cdot C + B \oplus C
        \end{split}
    \end{equation}

    The resulting visualisation in Fig. \ref{fig:rule110} exhibits similar
    characteristic of self-organisation, however it is hard to say whether it
    shows self-similarity as well.

    \begin{figure}[h]
        \begin{center}
            \includegraphics[width=4in]{rule110.png}
            \caption{Rule 110 generated using Eq. \ref{eq:rule110}}
            \label{fig:rule110}
        \end{center}
    \end{figure}

    \subsection{Elementary cellular automata}
    Elementary cellular automata are the simplest class of one-dimensional
    cellular automata; each cell may be in one of two states, 0 or 1, and rule
    sets only take into account the cell itself and its immediate neighbours
    (the nearest cell left and right of the cell being computed). Since there
    are a total of $2^3=8$ combinations of values for a neighbourhood of cells,
    and each of these neighbourhoods causes a cell to move to one of two states
    in the next generation, this means there are $2^8=256$ elementary cellular
    automata. Their evolution is often illustrated by each progressive ``new
    line'' showing the next generation of the automaton, as caused by applying
    the rule set to the current generation.

    Two technically different automata may give rise to similar patterns, and
    indeed many automata are equivalent to each other thanks to simple
    transformations of theirunerlying geometry. The first of these
    transformations is reflection in the vertical axis; the result of applying
    this is known as the mirror rule \cite{elementary}. For example, take Rule
    30. Each time some neighbourhood produces a live cell in Rule 30, let's take
    this neighbourhood and reflect it in the vertical plane, giving the table
    below:

    \begin{table}[h]
        \centering
        \begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
            111 & 110 & 101 & 100 & 011 & 010 & 001 & 000   \\
            \hline
            0 & 0 & 0 & 1 & 1 & 1 & 1 & 0   \\
            \hline
            0 & 1 & 0 & 1 & 0 & 1 & 1 & 0   \\
            \hline
        \end{tabular}
    \end{table}

    The first line of the table shows the ruleset for Rule 30, while the line
    below gives that of our new ruleset, made by applying the mirror rule to
    Rule 30. Thus we get Rule 30's mirror rule Rule $01010110_{\text{bin}}$ =
    Rule $86_{\text{dec}}$. Rules which generate the same ruleset under
    mirroring are called amphichiral, and of the 256 elementary cellular
    automata, 64 are so.

    The second of these transformations is to exchange the roles of 0s and 1s in
    the ruleset definition, and doing so is applying the complement rule
    \cite{elementary}.  Again applying this to Rule 30, we get the table below
    (note that the roles of 0s and 1s are exchanged in both the top and bottom
    rows):

    \begin{table}[h]
        \centering
        \begin{tabular}{|c|c|c|c|c|c|c|c|}\hline
            000 & 001 & 010 & 011 & 100 & 101 & 110 & 111   \\
            \hline
            1 & 1 & 1 & 0 & 0 & 0 & 0 & 1   \\
            \hline
        \end{tabular}
    \end{table}

    So we arrive at Rule $10000111_{\text{bin}} = 135_{\text{dec}}$ as the
    complement of Rule 30. There are 16 rules which are the same as their
    complement.
    Both of these rules can be applied at once to a ruleset, and the result is
    Rule 149 when applied to Rule 30. There are 16 rules which are the same as
    their mirrored complementary rules. There are 88 rules which are
    inequivalent under these transformations \cite{elementary}.

    \subsection{Wolfram classes}
    Wolfram gives the following classes for categorising the behaviours of
    different automata:

    \begin{itemize}
        \item Class One: rapidly converge to a uniform state e.g. Rule 0, 232
        \item Class Two: rapidly converge to a repetitive or stable state e.g.
            Rule 250
        \item Class Three: appear to remain in a random state e.g. Rule 150, 182
        \item Class Four: form areas of repetitive or stable states, but also
            structures which interact in complicated ways e.g. Rule 54
    \end{itemize}

    Two of these 256 rules are particularly special, namely Rule 54 and 110,
    both of which are Class 4 elementary cellular automata. Rule 110 has been
    proven to be universal, or Turing complete, meaning it is theorectically
    able to simulate any function. Meanwhile, it is so far unknown whether Rule
    54 is capable of universal computation -- interacting structures form, but
    it remains to been seen whether structures useful for computation can be
    formed. Computational universality is discussed more in-depth later on.

    \subsection{Two dimensional cellular automata}
    Two dimensional cellular automata become a bit more interesting than their
    dimensionally-deficient brethren. Rather than having a single row of cells
    to consider at each generation, now a two-dimensional plane of cells is used
    to compute the evolution of the automaton. Whereas elementary automata
    looked only at the neighbourhood of $a_{i-1}, a_i,$ and $a_{i+1}$ when
    applying the ruleset, two-dimensional automata have a wider choice. Two of
    the most common types of neighbourhood to consider are the Moore
    neighbourhood and the von Neumann neighbourhood. The former includes the 8
    cells directly surrounding the current cell being considered, while the
    latter only focuses on the four orthogonally ajacent cells. Ocassionally,
    one may use the extended von Neumann neighbourhood.

    \subsection{Life-like cellular automata}
    \subsection{Cyclic cellular automata}

\section{Examples}
\section{Applications}
\section{Computational Universality}
\section{Firing Squad Synchronisation Problem}
\section{Majority Problem}

\bibliography{bibliography}

\end{document}